---
title: "Mfg wage premium"
author: "Daniel Perez"
date: "3/15/2022"
output: html_document
---


```{r Libraries}
library(tidyverse)
library(epiextractr)
library(here)
library(labelled)
library(janitor)
library(broom)
```

```{r Load CPI data}
cpi <- read_csv(here("input/cpi_annual.csv")) 
cpi_base <- cpi$cpiurs[cpi$year == 2021] 

```

```{r Load CPS data}
#wbhao ORG 1989 - present
#selfinc 1989 - present

cpsorg <- load_org(1979:2021, year, month, orgwgt, lfstat, emp, selfemp, selfinc, manuf, wage, age, female, educ, gradeatn, wbho, wbhao, region) %>% 
  filter(age>=16, emp==1) %>% 
  #merge CPI to CPS data
  left_join(cpi, by='year') %>% 
  mutate(wgt = orgwgt/12) %>% 
  #inflation adjust wages to $2021
  mutate(realwage = wage * (cpi_base/cpiurs)) %>% 
  mutate(lnwage = log(realwage)) %>% 
  mutate(agebin = case_when(between(age,16,24) ~ 1, #create age bins
                            between(age,25,34) ~ 2,
                            between(age,35,44) ~ 3,
                            between(age,45,54) ~ 4,
                            between(age,55,64) ~ 5,
                            age>=65 ~ 6)) %>% 
  mutate(not_selfemp = ifelse(selfemp==0 | is.na(selfemp), yes=1, no=0),
         not_selfinc = ifelse(selfinc==0 | is.na(selfinc), yes=1, no=0)) %>% 
  set_value_labels(agebin = c("16-24"=1, "25-34"=2, "35-44"=3, "45-54"=4, "55-64"=5, "65&up"=6))  

```


```{r Average wages - all vs mfg}

### Disaggregate manuf from all workers

#removed selfinc from 1989-present
avg_wages <- cpsorg %>% 
filter(not_selfemp==1, not_selfinc==1) %>% 
group_by(year, manuf) %>% 
summarize(avg_wage = weighted.mean(realwage, w=wgt, na.rm=TRUE),
          n=n(),
          employment = sum(wgt)) %>% 
  mutate(manuf = to_factor(manuf)) %>% 
  pivot_wider(id_cols = year, names_from = manuf, values_from = c(avg_wage, n, employment)) %>% 
  clean_names() %>% 
  mutate(pct_diff_avgwage = (avg_wage_manufacturing/avg_wage_not_in_manufacturing)-1) %>% 
  write_csv(here(paste0("output/avg_wages_",format(Sys.time(), "%d-%b-%Y %H.%M"),".csv")))

#average wages for BA plus
avg_wages_collegeplus <- cpsorg %>% 
filter(not_selfemp==1, not_selfinc==1, educ>=4) %>% 
group_by(year, manuf) %>% 
summarize(avg_wage = weighted.mean(realwage, w=wgt, na.rm=TRUE),
          n=n(),
          employment = sum(wgt)) %>% 
  mutate(manuf = to_factor(manuf)) %>% 
  pivot_wider(id_cols = year, names_from = manuf, values_from = c(avg_wage, n, employment)) %>% 
  clean_names() %>% 
  mutate(pct_diff_avgwage = (avg_wage_manufacturing/avg_wage_not_in_manufacturing)-1)
  
#average wages for less than BA
avg_wages_somecoll <- cpsorg %>% 
filter(not_selfemp==1, not_selfinc==1, educ<4) %>% 
group_by(year, manuf) %>% 
summarize(avg_wage = weighted.mean(realwage, w=wgt, na.rm=TRUE),
          n=n(),
          employment = sum(wgt)) %>% 
  mutate(manuf = to_factor(manuf)) %>% 
  pivot_wider(id_cols = year, names_from = manuf, values_from = c(avg_wage, n, employment)) %>% 
  clean_names() %>% 
  mutate(pct_diff_avgwage = (avg_wage_manufacturing/avg_wage_not_in_manufacturing)-1)

```

No union dummy

I'd suggest adding age dummies, say 16-24, 25-34, 35-54, 55+. 

On Fri, Mar 11, 2022 at 2:28 PM L Josh Bivens <lbivens@epi.org> wrote:

I’d like a few estimates of the “manufacturing wage premium”.
 
1. One, just raw – average wage of manufacturing workers relative to average wage in all other industries (combined). Basically looking for (manufacturing wage/all other wages) - 1.
2. Second, regression-adjusted – regress hourly wages on 5 education category dummies, gender dummies, race/ethnicity dummies, region dummies, and a dummy indicating whether they work in manufacturing. That coefficient on manufacturing should range from .03 to .15. I don’t think I want to control for unionization for this one – cc’ing John so he can disagree with this if he thinks that’s wrong, and, happy to elaborate why, but, for now (unless we flip due to objections from John), we’ll run with no union control. The manufacturing wage premium is just the coefficient on the manufacturing dummy variable.
3. Third, regression-adjusted, but, just for workers with less than a 4-year college degree. So, regression-adjusted manufacturing premium from above, but, with sample exclusions to just include non-college workers.
4. Fourth, regression-adjusted, but, for workers with a 4-year degree and above. 
5. Would like all of those for as many years as microdata makes possible (73? 79?).


```{r Regressions mfg premium}

#dataframe for our regressions
  #Turn variables into factors for our categorical regression

wage_regs <- cpsorg %>% 
  filter(not_selfemp==1, not_selfinc==1) %>%
  mutate(across(educ|female|agebin|region|wbho|manuf, ~factor(.x))) %>% 
  group_by(year) %>% 
  #model for our regression
  summarize(lm(lnwage ~ manuf + educ + female + wbho + agebin + region) %>%  tidy()) %>% 
  filter(term=='manuf1') %>% 
  write_csv(here(paste0("output/mfg_wage_prem_",format(Sys.time(), "%d-%b-%Y %H.%M"),".csv")))

wage_regs_somecollege <- cpsorg %>% 
  filter(not_selfemp==1, not_selfinc==1, educ<4) %>%
  mutate(across(female|agebin|educ|region|wbho|manuf, ~factor(.x))) %>%
  #create dummy educ variable where lths = 0 and hs = 1
  group_by(year) %>% 
  #model for our regression
  summarize(lm(lnwage ~ manuf + female + wbho + agebin + educ + region) %>%  tidy()) %>% 
  filter(term=='manuf1') %>% 
  write_csv(here(paste0("output/wage_regs_somecollege_",format(Sys.time(), "%d-%b-%Y %H.%M"),".csv")))


wage_regs_4yearplus <- cpsorg %>% 
  filter(not_selfemp==1, not_selfinc==1, educ>=4) %>%
  mutate(across(female|agebin|region|educ|wbho|manuf, ~factor(.x))) %>%
  group_by(year) %>% 
  #model for our regression
  summarize(lm(lnwage ~ manuf + female + wbho + agebin + educ + region) %>%  tidy()) %>% 
  filter(term=='manuf1') %>% 
  write_csv(here(paste0("output/wage_regs_baplus_",format(Sys.time(), "%d-%b-%Y %H.%M"),".csv")))

```



